Modelling last glacial cycle ice dynamics in the Alps

### 2017-07-19 articles/cycle/alps-cycle.tex
Add palaeo-climate sensitivity first shot.

    \begin{itemize}
    \item All yield glaciation during MIS~4 and 2 (Fig.~\ref{fig:timeseries}).
    \item But otherwise very different results (Fig.~\ref{fig:timeseries}).
    \end{itemize}

@@

    \begin{itemize}
    \item Two yield two large MIS~4 ice extent (Fig.~\ref{fig:footprints}).
    \item Only EPICA yield realistic MIS~4 (Fig.~\ref{fig:footprints}).
    \end{itemize}

>>>>

    For the three palaeo-temperature records and the two palaeo-precipitation
    parametrizations used, the model yield significant ice volume build-up
    during MIS~4 and 2 (Fig.~\ref{fig:timeseries}). All simulations also
    yield important glaciations during MIS~5 and 3, but timing and amplitude
    varies a lot between the different forcing. All simulations overestimate
    ice cover during the Younger Dryas. This might be due partly to the 2\,km
    resolution, except for GRIP which largely overestimates Younger Dryas ice
    cover.

    Even more importantly, all simulations yield very strong ice volume
    variability which is the result of multiple glacier advance and retreats
    over the foreland, many more than two. Palaeo-precipitation reductions
    partly results in a smoother modelled ice volume time series and less
    extreme variations, but variability remains high. Ice volume variability
    is the smallest for the EPICA palaeo-temperature record, which has the
    least temperature variability.

@@

    To select an optimal climate forcing, we look at the best known alpine
    glaciations during MIS~2 and 4. During MIS~2, all forcing yield ice extent
    of the same order of magnitude (Fig.~\ref{fig:footprints}). This is
    because the temperature records have been scaled to resul in similar
    ice covered area for the Rhine Glacier piedmont lobe. Looking at ice cover
    on other parts, all simulations tend to overestimate ice extent in the
    eastern part of the model domain, and to underestimate it in the western
    part, as compared to the reconstructed LGM margin. Regarding the western
    part, there are discussions though, as if the Lyon Lobe actually dates from
    MIS~2 or from an older glaciation. Regarding the eastern part, the MIS~2
    glaciation has been mapped in detail and dated (?), so that there is no
    doubt that our model overestimate ice cover in this region. This could
    indicate that during LGM the east-west precipitation gradient over the Alps
    was lower than today, or that temperature depression in the east was lower
    than in the west.

    There are spatial differences though: MD01-2444, and to a greater extent
    GRIP, forcings, tend to overestimate ice cover in all peripherical ranges:
    the Vosges, Black Forest, Bavarian Forest, Dinaric Alps and Pennine. This
    is because these records have a higher temperature variability so that
    after scaling they comport brief periods of cold climate, not long enough
    to develop the alpine ice cap, but long enough to build up ice on the
    peripherical ranges.

    The extent of ice cover during MIS~4 shows more sensitivity to the choice
    of palaeo-climate forcing. Using the GRIP and MD01-2444, glaciers extend
    well beyound the reach of documented moraines and glacial erratic terrain.
    Because this extent corresponds to a short-term cold climate event,
    palaeo-precipitation reductions help to greatly reduce this excessive
    modelled ice cover, yet in both cases, modelled ice extent and volume
    during MIS~4 remains higher than during MIS~2. EPICA forcing, on the other
    hand, yield a MIS~4 glaciation that is only slightly less expensive than
    the MIS~2, with only little sensitivity of the glaciated area to
    palaeo-precipitation reductions.

    Based on the above consideration on timing of the LGM ice extent and
    ice extent during MIS~4, we chose EPICA as our optimal temperature record
    for the rest of this paper. As a conservative approach in regard to quick
    ice volume fluctuations, we choose to include palaeo-precipitation
    correction in the following 1\,km-resolution simulation.
